### 2011-02-09 paper/learningMethods.tex
Edits to learningMethods, took out SVM for now

\edit{make sure this text flows into the table}

@@

\noindent Here the fixed constants that determine the crouched
position are $C_I$ for inner motors, $C_O$ for outer motors, and $C_C$
for the center hip motor.  The nine subscripted functions correspond
to the nine labeled motors in \figref{topdown.png}.

@@

If we choose a value for the five dimensional parameter
$\vec{\theta}$, then a given physical trial gives us one measurement
of the fitness of that parameter vector, or $f(\vec{\theta})$.  Two
things make learning difficult.  First, each evaluation of
$f(\vec{\theta})$ is expensive, taking 15-20 seconds of time on
average.  Second, the fitness returned by such evaluations has proved
to be very noisy, with the standard deviation of the noise often being
roughly equivalent to the size of the measurement.

@@

We evaluated a total of seven subtly or not so subtly different
learning methods for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial.  In
general this is done by selecting a random starting vector within the
allowable domain.  However, as discussed in
\secref{fitnessEvaluation}, we decided to evaluate the different
methods starting at one of several common starting points to reduce
one possible contribution of noise.

@@

\item \emph{Uniform random hill climbing}: This method begins by
  selecting a single random parameter vector.  Subsequent iterations
  generate a neighbor by randomly choosing one parameter to adjust and
  replacing it with a new value chosen with uniform probability in the
  allowable range for that parameter. The neighbor is evaluated by
  running the robot with the newly chosen parameters. If this neighbor
  results in a longer distance walked than the previous best gait, it
  is saved as the new best gait. The process is then repeated, always
  starting with the best gait.

\item \emph{Gaussian random hill climbing}: This method works
  similarly to Uniform random hill climbing, except neighbors are
  generated by adding random Gaussian noise to the current best gait.
  This results in all parameters being changed at once, but the
  resulting vector is always fairly close to the previous best gait.
  We used independently selected noise in each dimension, scaled such
  that the standard deviation of the noise was 5\% of the range of
  that dimension.

\item \emph{N-dimensional policy gradient descent}: In \cite{kohl},
  Kohl and Stone document a method for local gradient ascent for gait
  learning with noisy fitness evaluations, and we have included this
  method in our evaluation.  This strategy explicitly estimates the
  gradient for the objective function. It does this by first
  evaluating \emph{n} randomly generated parameter vectors near the
  initial vector, each dimension of these vectors being perturbed by
  either $-\epsilon$, $0$, or $\epsilon$. Then, for each dimension, it
  groups vectors into three groups: $-\epsilon$, $0$, and $\epsilon$.
  The gradient along this dimension is then estimated as the average
  score for the $\epsilon$ group minus the average score for the
  $-\epsilon$ group. Finally, the method creates a new vector by
  changing all parameters by a fixed-size step in the direction of the
  gradient.

\item \emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method \cite{nm} creates an initial simplex with $d+1$ vertices,
  where $d$ is the dimension of the parameter space. The initial
  parameter vector is stored as the first vertex and the other five
  vertices are created by adding to one dimension at a time one tenth
  of the allowable range for that parameter. It then tests the fitness
  of each vertex and based on these fitnesses, it reflects the worst
  point over the centroid in an attempt to improve it.  In general,
  the worst vertex is reflected; however, to prevent cycles and
  becoming stuck in local minima, several other rules are used.  If
  the reflected point is better than the second worst point and worse
  than the best point, then the reflected point replaces the worst. If
  the reflected point is better than the best point, the simplex is
  expanded in the direction of the reflected point. The better of the
  reflected and the expanded point replaces the worst point. If the
  reflected point is worse than the second worst point, then the
  simplex is contracted away from the reflected point. If the
  contracted point is better than the reflected point, the contracted
  point replaces the worst point. If the contracted point is worse
  than the reflected point, the entire simplex is shrunk \cite{nm}.

\item \emph{Linear regression}: To initialize, this method chooses and
  evaluates five random parameter vectors. It then fits a linear model
  from parameter vector to fitness. In a loop, the method chooses and
  evaluates a new parameter vector generated by taking a fixed-size
  step in the direction of the gradient for each parameter, and fits a
  new linear model to all vectors evaluated so far, choosing the model
  to minimize the sum of squared errors.

\item \emph{SVM regression}: Similarly to linear regression, this
  model starts with several random vectors, but this time they are
  chosen in a small neighborhood about some initial random vector.
  These vectors (generally 8) are evaluated, and a support vector
  regression model is fit to the observed fitnesses.  To choose the
  next vector for evaluation, we randomly generate some number
  (typically 100) of vectors in the neighborhood of the best observed
  gait, and select for evaluation the vector with the best predicted
  performance.  We suspected that if we always chose the best
  predicted point out of 100, we may end up progressing along a narrow
  subspace, prohibiting learning of the true local fitness function.
  Put another way, we would always choose exploitation of knowledge
  vs. exploration of the space.  To address this concern, we added a
  parameter dubbed \code{bumpBy} that added noise to the final
  selected point before it was evaluated.

  Such a method naturally has many tunable parameters, and we
  endeavored to select these parameters by tuning the method in
  simulation.  To estimate the performance of the algorithm, we ran it
  against a simulation with a known optimum.  The simulated function
  was in the same five dimensional parameter space, and simply
  returned a fitness determined as the height of a Gaussian with a
  random mean.  The width of the Gaussian in each dimension was 20\%
  of the range of each dimension, and the maximum value at the peak
  was 100.  \figref{svm_sim_results} shows the learning results on
  this simulated model using the ultimately selected SVM parameters.
  Interestingly, a non-zero value of \code{bumpBy} resulted in better
  learning than noise free (exploration free) learning.

  \edit{JMC: This next paragraph sounds like it belongs in Results?}
  Ultimately, however, the version of SVM tuned for simulation still
  did not show competitive performance on the real robot.  We tried
  tuning some parameters on the real robot, but after some amount of
  tuning, the method still exhibited too little exploration and easily
  became stuck in local minima.

>>>>

@@

\noindent Here the fixed constants that determine the crouched
position are $C_I = 800$ for inner motors, $C_O = 40$ for outer
motors, and $C_C = 512$ for the center hip motor.  The nine functions
correspond to the nine motors in the same order as listed in
\figref{topdown.png}.

@@

If we choose a value for the five dimensional parameter
$\vec{\theta}$, then a given physical trial gives us one measurement
of the fitness $f(\vec{\theta})$ of that parameter vector.  Two
things make learning difficult.  First, each evaluation of
$f(\vec{\theta})$ is expensive, taking 15-20 seconds of time on
average.  Second, the fitness returned by such evaluations has proved
to be very noisy, with the standard deviation of the noise often being
roughly equivalent to the size of the measurement.

@@

We evaluated a total of seven subtly or not so subtly different
learning methods for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial.  In
general this is done by selecting a random starting vector within the
allowable domain.  However, to most directly compare learning methods
by their ability to improve on the initially chosen $\vec{\theta}$, we
decided to evaluate the different methods by starting them at the same
initial $\vec{\theta}$ vectors.

@@

\item \emph{Uniform random hill climbing}: This method always starts
  with the current best gait, and then chooses next $\vec{\theta}$ by
  randomly choosing one parameter to adjust and replacing it with a
  new value chosen with uniform probability in the allowable range for
  that parameter. This new point is evaluated, and if it results in a
  longer distance walked than the previous best gait, it is saved as
  the new best gait. The process is then repeated, always starting
  with the best gait.

\item \emph{Gaussian random hill climbing}: This method works
  similarly to Uniform random hill climbing, except the next
  $\vec{\theta}$ is generated by adding random Gaussian noise to the
  current best gait.  This results in all parameters being changed at
  once, but the resulting vector is always fairly close to the
  previous best gait.  We used independently selected noise in each
  dimension, scaled such that the standard deviation of the noise was
  5\% of the range of that dimension.

\item \emph{N-dimensional policy gradient ascent}: We implemented Kohl
  and Stone's \cite{kohl} method for local gradient ascent for gait
  learning with noisy fitness evaluations. This strategy explicitly
  estimates the gradient of the objective function. It does this by
  first generating $n$ parameter vectors near the initial vector by
  perturbing each dimension of each vector randomly by either
  $-\epsilon$, $0$, or $\epsilon$. Then each vector is run on the
  robot, and for each dimension we group the results into three
  groups: $-\epsilon$, $0$, and $\epsilon$.  The gradient along this
  dimension is then estimated as the average score for the $\epsilon$
  group minus the average score for the $-\epsilon$ group. Finally,
  the method creates the next $\vec{\theta}$ by changing all
  parameters by a fixed-size step in the direction of the gradient.
  For this study we used values of $\epsilon$ equal to 5\% of the
  allowable range in each dimension (ranges listed in
  \tabref{parameters}), and a step size scaled such that if all dimensions
  were in the range [0,~1], the norm of the step size would be 0.1.

\item \emph{Nelder-Mead simplex method}: The Nelder-Mead simplex
  method \cite{nm} creates an initial simplex with $d+1$ vertices,
  where $d$ is the dimension of the parameter space. The initial
  parameter vector is stored as the first vertex and the other five
  vertices are created by adding to one dimension at a time one tenth
  of the allowable range for that parameter. It then tests the fitness
  of each vertex, and, in general, it reflects the worst point over
  the centroid in an attempt to improve it.  However, to prevent
  cycles and becoming stuck in local minima, several other rules are
  used.  If the reflected point is better than the second worst point
  and worse than the best point, then the reflected point replaces the
  worst. If the reflected point is better than the best point, the
  simplex is expanded in the direction of the reflected point. The
  better of the reflected and the expanded point replaces the worst
  point. If the reflected point is worse than the second worst point,
  then the simplex is contracted away from the reflected point. If the
  contracted point is better than the reflected point, the contracted
  point replaces the worst point. If the contracted point is worse
  than the reflected point, the entire simplex is shrunk \cite{nm}.

\item \emph{Linear regression}: To initialize, this method chooses and
  evaluates five random parameter vectors. It then fits a linear model
  from parameter vector to fitness. In a loop, the method chooses and
  evaluates a new parameter vector generated by taking a fixed-size
  step in the direction of the gradient for each parameter, and fits a
  new linear model to all vectors evaluated so far, choosing the model
  to minimize the sum of squared errors. Step size is the same as in
  \emph{N-dimensional policy gradient ascent}.

%\item \emph{SVM regression}: Similarly to linear regression, this
%  model starts with several random vectors, but this time they are
%  chosen in a small neighborhood about some initial random vector.
%  These vectors (generally 8) are evaluated, and a support vector
%  regression model is fit to the observed fitnesses.  To choose the
%  next vector for evaluation, we randomly generate 100 vectors in the
%  neighborhood of the best observed gait, and select for evaluation
%  the vector with the best predicted performance in the learned model.
%  We suspected that if we always chose the best predicted point out of
%  100, we may end up progressing along a narrow subspace, prohibiting
%  learning of the true local fitness function.  Put another way, we
%  would always choose exploitation of knowledge vs. exploration of the
%  space.  To address this concern, we added a small amount of noise to
%  the new chosen $\vec{\theta}$ before continuing with the next phase
%  of evaluation and modeling.
%
%  Such a method naturally has many tunable parameters, and we
%  endeavored to select these parameters by tuning the method in
%  simulation.  To estimate the performance of the algorithm, we ran it
%  against a simulation with a known optimum.  The simulated function
%  was in the same five dimensional parameter space, and simply
%  returned a fitness determined as the height of a Gaussian with a
%  random mean.  The width of the Gaussian in each dimension was 20\%
%  of the range of each dimension, and the maximum value at the peak
%  was 100.  \figref{svm_sim_results} shows the learning results on
%  this simulated model using the ultimately selected SVM parameters.
%  Interestingly, a non-zero value of \code{bumpBy} resulted in better
%  learning than noise free (exploration free) learning.
%
%  \edit{JMC: This next paragraph sounds like it belongs in Results?}
%  Ultimately, however, the version of SVM tuned for simulation still
%  did not show competitive performance on the real robot.  We tried
%  tuning some parameters on the real robot, but after some amount of
%  tuning, the method still exhibited too little exploration and easily
%  became stuck in local minima.
